---
title: "Project 2"
output: html_document
date: "2024-11-10"
author: "Tade Scheid"
---

## Libraries
```{r}
library(ggplot2)
library(janitor)
library(ggrepel)
```

## Imported Data Set
```{r}
wnba.stats <- read.csv("data/wnba-player-stats.csv")
clean.wnbastats <-remove_empty(wnba.stats)
```

## New Visual
```{r, fig.width=8, fig.height=5}
plot.df <- clean.wnbastats[clean.wnbastats$Pos %in% c("C", "C-F", "F", "F-C", "F-G", "G", "G-F" ), ]
plot.df <- plot.df[order(plot.df$Age), ]
mean_FTr <- function(age, pos){mean(plot.df[plot.df[ , "Age"] == age & plot.df[ , "Pos"] == pos, "FTr"])}

wnba_df <- data.frame("Pos" = character(0), "Age" = numeric(0), "FTr" = numeric(0))
for(wnba_age in unique(plot.df$Age)){
  for(wnba_pos in unique(plot.df$Pos)){
    newrow = c(wnba_pos, wnba_age, mean_FTr(wnba_age, wnba_pos)) 
    wnba_df <- rbind(wnba_df, newrow)
  }
}

plot(wnba_df$X..1., wnba_df$X.0.121., type = "n", 
     xlab = "Age", ylab = "Average Free Throw Rate", 
    xlim = c(18,50), ylim = c(0, 1), 
     main = "Average Free Throw Rate over Age by Position for WNBA 2019 Season")

color_lines <- rainbow(length(unique(wnba_df$X.G.)))

for (i in 1:length(unique(wnba_df$X.G.))) {
  position = unique(wnba_df$X.G.)[i]
  lines(wnba_df[wnba_df$X.G. == position, "X..1."], 
        wnba_df[wnba_df$X.G. == position, "X.0.121."], 
        col = color_lines[i], lwd = 2)
}

label_dict <- c("C" = "Center", "C-F" = "Center-Forward", "F" = "Forward", "F-C" = "Forward-Center", "F-G" = "Forward-Guard", 
                "G" = "Guard", "G-F" = "Guard-Forward")

legend("topright", legend = label_dict[unique(wnba_df$X.G.)], col = color_lines, lty = 1, lwd = 2)

```

The graph plots Average Free Throw Rate of each position over Age for the 2019 WNBA Season. The positions it represents by each colored line are Guard, Center, Forward-Center, Forward-Guard, Forward, Guard-Forward, and Center-Forward which are specified in the legend. The graph demonstrates how Center-Forwards have the highest average free throw rates which increase over time. Forward-Guards have the lowest average free throw rates which stays pretty consistant over time. 


## Changes Made
I changed the visualization I made by plotting Age vs Average Free Throw Rate on a line graph instead of Age vs Player Efficacy Rating on a scatterplot. I decided to include the element of positions from my second graph in Project 1 by making each line graphed represent a different position. The colored lines were also instead of labeling the outliers with the individual player's name.

## Sources
https://web.itu.edu.tr/~tokerem/The_Book_of_R.pdf
https://github.com/fivethirtyeight/WNBA-stats/blob/master/wnba-player-stats.csv

